### 2018-06-06 ap_association_for_proportional_fairness_in_mu-mimo_multi-rate_wlans/AP_Association_Analysis.tex

% AP Association Analysis
\section{Analysis}
	After solving the convex program, we can obtain $b_j^* = \sum_{i \in A} b_{ij}^*, \forall j \in U.$
	For each user $j$, we associate user $j$ with AP $i'$ such that $b_{i'j}^*$ is maximum in $\{b_{ij}^*|\forall i \in A \}.$ 
	Let $$\hat{b}_j = b_{i'j}^*, \forall j \in U.$$
	Then, $$\hat{b}_j \ge \frac{b_j^*}{|A|}.$$
	We know $$\sum_{j \in U} \log{b_j} = \sum_{i \in A} \sum_{j \in U_i} \log{b_j} \ge \sum_{i \in A} \sum_{j \in U_i} \log{\hat{b}_j} = \sum_{j \in U} \log{\hat{b}_j}.$$
	Then, $$\sum_{j \in U} \log{\hat{b}_j} \ge \sum_{j \in U} \log{\frac{b_j^*}{|A|}} $$ $$ = \sum_{j \in U} (\log{b_j^*}-\log{|A|}) = \sum_{j \in U} \log{b_j^*} - |U|\log{|A|} \ge OPT - |U|\log{|A|}.$$