### 2018-06-06 ap_association_for_proportional_fairness_in_mu-mimo_multi-rate_wlans/AP_Assoication_for_Proportional_Fairness_in_MU-MIMO_Multi-rate_WLANs.tex

%
\documentclass[a4paper, 12pt]{report}
% packages
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{algorithm} 
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[toc]{appendix}
% style
\topmargin 0cm \oddsidemargin 0.5cm \headsep 1.5cm \headheight 0.5cm
\textheight 21.5cm \textwidth 15.2cm \marginparwidth 0.5cm
\marginparsep 0cm
\floatsep 0.8cm     %頁之上端與下端圖表與圖表之間的間距
\textfloatsep 0.8cm %頁之上端與下端圖表與文字之間的間距
\intextsep 1.0cm    %指定為 h 的圖表與文字之間的間距
% new command
\newcommand{\doublespace}{\addtolength{\baselineskip}{.6\baselineskip}}

% title and author
\title{AP Association for Proportional Fairness in MU-MIMO Multi-rate WLANs}
\author{YuShu Chen}
%
\begin{document}
	\begin{titlepage}
%		\maketitle
		\begin{center}
			%\vspace{2.5cm}
			\Large{\bf{AP Association for Proportional Fairness in MU-MIMO Multi-rate WLANs\\}}
			\vspace{3cm}
			Department of Computer Science\\
			National Tsing Hua University\\
			\vspace{2cm}
			%(Draft)\\
			\vspace{4.5cm}
			\parbox[t]{7cm}{
				%實際寬度自行再調整
				Advisor: Dr. Ming-Jer~Tsai\\
				Student: Yu-Shu~Chen\\
			} \\
			\vspace{5cm}
			July 2015\\
		\end{center}
	\end{titlepage}
	\pagenumbering{roman}
	\doublespace
	\chapter*{Abstract}
		\input{Abstract.tex}
	% table of content
	\addcontentsline{toc}{chapter}{Contents}
	\tableofcontents
	\cleardoublepage

	\addcontentsline{toc}{chapter}{List of Figures}
	\listoffigures
	\cleardoublepage

	\addcontentsline{toc}{chapter}{List of Tables}
	\listoftables
	\cleardoublepage

	\pagenumbering{arabic}
	\chapter{Introduction and Related Work}
		In this chapter, we introduce the AP association problem in addition to background knowledge and show the related work.
		\input{Introduction.tex}
		\input{Related_Work.tex}
	\chapter{Bandwidth Allocation}
		In this chapter, we describe how to allocate bandwidth for proportional fairness to each user associating with the same AP.
		\input{Bandwidth_Allocation.tex}
	\chapter{AP Association}
		In this chapter, we describe the AP association problem AAPFM and propose an algorithm called CVXP to solve it. We will analysis CVXP and show the hardness of this problem.
		\input{AP_Association.tex}
		\input{AP_Association_Analysis.tex}
		\input{AP_Association_Hardness.tex}
	\chapter{Performance Evaluations}
		In this chapter, we build a bit rate table for simulations and compare the result of our algorithms CVXP to a naive algorithm HDR in each simulation.
		\input{Bit_Rate_Table.tex}
%		\input{Evaluation_for_AP_Association.tex}
		\input{Evaluation_for_AP_Association_2.tex}
	\chapter{Conclusion}
		\input{Conclusion.tex}
%	\input{Bibliography.tex}
	\bibliographystyle{IEEEtran}
	\bibliography{references}
	\begin{appendices}
		\chapter{Details of Simulations}
		\label{apendix:dos}
		\input{Details_of_Simulations}
	\end{appendices}
\end{document}