### 2018-06-06 ap_association_for_proportional_fairness_in_mu-mimo_multi-rate_wlans/Details_of_Simulations.tex

% Details of Simulations

\subsubsection{SIMUL1-1-1}
	Fig. \ref{figure:simul1_1_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul1_1_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul1_1_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL1-1-2}
	Fig. \ref{figure:simul1_1_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul1_1_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul1_1_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL1-2-1}
	Fig. \ref{figure:simul1_2_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul1_2_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul1_2_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL1-2-2}
	Fig. \ref{figure:simul1_2_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul1_2_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul1_2_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL1-3-1}
	Fig. \ref{figure:simul1_3_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul1_3_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul1_3_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL1-3-2}
	Fig. \ref{figure:simul1_3_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul1_3_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul1_3_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL1-4}
	Fig. \ref{figure:simul1_4_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul1_4_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul1_4_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL2-1-1}
	Fig. \ref{figure:simul2_1_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul2_1_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul2_1_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL2-1-2}
	Fig. \ref{figure:simul2_1_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul2_1_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul2_1_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL2-2-1}
	Fig. \ref{figure:simul2_2_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul2_2_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul2_2_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL2-2-2}
	Fig. \ref{figure:simul2_2_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul2_2_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul2_2_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL2-3-1}
	Fig. \ref{figure:simul2_3_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul2_3_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul2_3_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL2-3-2}
	Fig. \ref{figure:simul2_3_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul2_3_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul2_3_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL2-4}
	Fig. \ref{figure:simul2_4_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul2_4_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul2_4_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL3-1-1}
	Fig. \ref{figure:simul3_1_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul3_1_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul3_1_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL3-1-2}
	Fig. \ref{figure:simul3_1_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul3_1_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul3_1_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL3-2-1}
	Fig. \ref{figure:simul3_2_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul3_2_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul3_2_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL3-2-2}
	Fig. \ref{figure:simul3_2_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul3_2_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul3_2_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL3-3-1}
	Fig. \ref{figure:simul3_3_1_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul3_3_1_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul3_3_1_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL3-3-2}
	Fig. \ref{figure:simul3_3_2_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul3_3_2_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul3_3_2_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

\subsubsection{SIMUL3-4}
	Fig. \ref{figure:simul3_4_p} shows $10$ patterns of the distribution of positions of users in simulation. Fig. \ref{figure:simul3_4_ba} shows the sorted bandwidth allocated to users. Fig. \ref{figure:simul3_4_u} shows the comparison of the sum of the utility function and the number of users obtaining more bandwidth between CVXP and HDR.

%=======
%\subsubsection{SIMUL1-1-1}
	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_01_position_v4.eps}
				\caption{pattern 1}
				\label{figure:simul1_1_1_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_02_position_v4.eps}
				\caption{pattern 2}
				\label{figure:simul1_1_1_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_03_position_v4.eps}
				\caption{pattern 3}
				\label{figure:simul1_1_1_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_04_position_v4.eps}
				\caption{pattern 4}
				\label{figure:simul1_1_1_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_05_position_v4.eps}
				\caption{pattern 5}
				\label{figure:simul1_1_1_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_06_position_v4.eps}
				\caption{pattern 6}
				\label{figure:simul1_1_1_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_07_position_v4.eps}
				\caption{pattern 7}
				\label{figure:simul1_1_1_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_08_position_v4.eps}
				\caption{pattern 8}
				\label{figure:simul1_1_1_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_09_position_v4.eps}
				\caption{pattern 9}
				\label{figure:simul1_1_1_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_10_position_v4.eps}
				\caption{pattern 10}
				\label{figure:simul1_1_1_j}
			\end{subfigure}
			\caption{Patterns of distribution of users in SIMUL1-1-1.}
			\label{figure:simul1_1_1_p}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul1_1_1_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul1_1_1_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul1_1_1_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul1_1_1_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul1_1_1_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul1_1_1_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul1_1_1_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul1_1_1_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul1_1_1_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp17_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul1_1_1_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL1-1-1.}
			\label{figure:simul1_1_1_ba}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=\textwidth]{fig/exp17_1-1-1_utility.eps}
				\caption{sum of utility function}
				\label{figure:simul1_1_1_u_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=\textwidth]{fig/exp17_1-1-1_bandwidth.eps}
				\caption{number of users}
				\label{figure:simul1_1_1_u_b}
			\end{subfigure}
			\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL1-1-1.}
			\label{figure:simul1_1_1_u}
		\end{center}
	\end{figure}

\clearpage

%\subsubsection{SIMUL1-1-2}
	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_01_position_v4.eps}
				\caption{pattern 1}
				\label{figure:simul1_1_2_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_02_position_v4.eps}
				\caption{pattern 2}
				\label{figure:simul1_1_2_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_03_position_v4.eps}
				\caption{pattern 3}
				\label{figure:simul1_1_2_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_04_position_v4.eps}
				\caption{pattern 4}
				\label{figure:simul1_1_2_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_05_position_v4.eps}
				\caption{pattern 5}
				\label{figure:simul1_1_2_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_06_position_v4.eps}
				\caption{pattern 6}
				\label{figure:simul1_1_2_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_07_position_v4.eps}
				\caption{pattern 7}
				\label{figure:simul1_1_2_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_08_position_v4.eps}
				\caption{pattern 8}
				\label{figure:simul1_1_2_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_09_position_v4.eps}
				\caption{pattern 9}
				\label{figure:simul1_1_2_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_10_position_v4.eps}
				\caption{pattern 10}
				\label{figure:simul1_1_2_j}
			\end{subfigure}
			\caption{Patterns of distribution of users in SIMUL1-1-2.}
			\label{figure:simul1_1_2_p}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul1_1_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul1_1_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul1_1_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul1_1_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul1_1_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul1_1_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul1_1_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul1_1_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul1_1_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp20_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul1_1_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL1-1-2.}
			\label{figure:simul1_1_2_ba}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp20_1-1-2_utility.eps}
				\caption{sum of utility function}
				\label{figure:simul1_1_2_u_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp20_1-1-2_bandwidth.eps}
				\caption{number of users}
				\label{figure:simul1_1_2_u_b}
			\end{subfigure}
			\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL1-1-2.}
			\label{figure:simul1_1_2_u}
		\end{center}
	\end{figure}

\clearpage

%\subsubsection{SIMUL1-2-1}
	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_01_position_v4.eps}
				\caption{pattern 1}
				\label{figure:simul1_2_1_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_02_position_v4.eps}
				\caption{pattern 2}
				\label{figure:simul1_2_1_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_03_position_v4.eps}
				\caption{pattern 3}
				\label{figure:simul1_2_1_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_04_position_v4.eps}
				\caption{pattern 4}
				\label{figure:simul1_2_1_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_05_position_v4.eps}
				\caption{pattern 5}
				\label{figure:simul1_2_1_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_06_position_v4.eps}
				\caption{pattern 6}
				\label{figure:simul1_2_1_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_07_position_v4.eps}
				\caption{pattern 7}
				\label{figure:simul1_2_1_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_08_position_v4.eps}
				\caption{pattern 8}
				\label{figure:simul1_2_1_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_09_position_v4.eps}
				\caption{pattern 9}
				\label{figure:simul1_2_1_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_10_position_v4.eps}
				\caption{pattern 10}
				\label{figure:simul1_2_1_j}
			\end{subfigure}
			\caption{Patterns of distribution of users in SIMUL1-2-1.}
			\label{figure:simul1_2_1_p}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul1_2_1_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul1_2_1_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul1_2_1_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul1_2_1_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul1_2_1_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul1_2_1_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul1_2_1_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul1_2_1_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul1_2_1_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp18_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul1_2_1_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL1-2-1.}
			\label{figure:simul1_2_1_ba}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp18_1-2-1_utility.eps}
				\caption{sum of utility function}
				\label{figure:simul1_2_1_u_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp18_1-2-1_bandwidth.eps}
				\caption{number of users}
				\label{figure:simul1_2_1_u_b}
			\end{subfigure}
			\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL1-2-1.}
			\label{figure:simul1_2_1_u}
		\end{center}
	\end{figure}

\clearpage

%\subsubsection{SIMUL1-2-2}
	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_01_position_v4.eps}
				\caption{pattern 1}
				\label{figure:simul1_2_2_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_02_position_v4.eps}
				\caption{pattern 2}
				\label{figure:simul1_2_2_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_03_position_v4.eps}
				\caption{pattern 3}
				\label{figure:simul1_2_2_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_04_position_v4.eps}
				\caption{pattern 4}
				\label{figure:simul1_2_2_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_05_position_v4.eps}
				\caption{pattern 5}
				\label{figure:simul1_2_2_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_06_position_v4.eps}
				\caption{pattern 6}
				\label{figure:simul1_2_2_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_07_position_v4.eps}
				\caption{pattern 7}
				\label{figure:simul1_2_2_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_08_position_v4.eps}
				\caption{pattern 8}
				\label{figure:simul1_2_2_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_09_position_v4.eps}
				\caption{pattern 9}
				\label{figure:simul1_2_2_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_10_position_v4.eps}
				\caption{pattern 10}
				\label{figure:simul1_2_2_j}
			\end{subfigure}
			\caption{Patterns of distribution of users in SIMUL1-2-2.}
			\label{figure:simul1_2_2_p}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul1_2_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul1_2_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul1_2_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul1_2_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul1_2_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul1_2_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul1_2_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul1_2_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul1_2_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp21_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul1_2_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL1-2-2.}
			\label{figure:simul1_2_2_ba}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp21_1-2-2_utility.eps}
				\caption{sum of utility function}
				\label{figure:simul1_2_2_u_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp21_1-2-2_bandwidth.eps}
				\caption{number of users}
				\label{figure:simul1_2_2_u_b}
			\end{subfigure}
			\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL1-2-2.}
			\label{figure:simul1_2_2_u}
		\end{center}
	\end{figure}

\clearpage

%\subsubsection{SIMUL1-3-1}
	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_01_position_v4.eps}
				\caption{pattern 1}
				\label{figure:simul1_3_1_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_02_position_v4.eps}
				\caption{pattern 2}
				\label{figure:simul1_3_1_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_03_position_v4.eps}
				\caption{pattern 3}
				\label{figure:simul1_3_1_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_04_position_v4.eps}
				\caption{pattern 4}
				\label{figure:simul1_3_1_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_05_position_v4.eps}
				\caption{pattern 5}
				\label{figure:simul1_3_1_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_06_position_v4.eps}
				\caption{pattern 6}
				\label{figure:simul1_3_1_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_07_position_v4.eps}
				\caption{pattern 7}
				\label{figure:simul1_3_1_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_08_position_v4.eps}
				\caption{pattern 8}
				\label{figure:simul1_3_1_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_09_position_v4.eps}
				\caption{pattern 9}
				\label{figure:simul1_3_1_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_10_position_v4.eps}
				\caption{pattern 10}
				\label{figure:simul1_3_1_j}
			\end{subfigure}
			\caption{Patterns of distribution of users in SIMUL1-3-1.}
			\label{figure:simul1_3_1_p}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul1_3_1_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul1_3_1_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul1_3_1_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul1_3_1_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul1_3_1_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul1_3_1_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul1_3_1_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul1_3_1_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul1_3_1_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp19_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul1_3_1_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL1-3-1.}
			\label{figure:simul1_3_1_ba}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp19_1-3-1_utility.eps}
				\caption{sum of utility function}
				\label{figure:simul1_3_1_u_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp19_1-3-1_bandwidth.eps}
				\caption{number of users}
				\label{figure:simul1_3_1_u_b}
			\end{subfigure}
			\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL1-3-1.}
			\label{figure:simul1_3_1_u}
		\end{center}
	\end{figure}

\clearpage

%\subsubsection{SIMUL1-3-2}
	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_01_position_v4.eps}
				\caption{pattern 1}
				\label{figure:simul1_3_2_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_02_position_v4.eps}
				\caption{pattern 2}
				\label{figure:simul1_3_2_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_03_position_v4.eps}
				\caption{pattern 3}
				\label{figure:simul1_3_2_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_04_position_v4.eps}
				\caption{pattern 4}
				\label{figure:simul1_3_2_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_05_position_v4.eps}
				\caption{pattern 5}
				\label{figure:simul1_3_2_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_06_position_v4.eps}
				\caption{pattern 6}
				\label{figure:simul1_3_2_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_07_position_v4.eps}
				\caption{pattern 7}
				\label{figure:simul1_3_2_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_08_position_v4.eps}
				\caption{pattern 8}
				\label{figure:simul1_3_2_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_09_position_v4.eps}
				\caption{pattern 9}
				\label{figure:simul1_3_2_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_10_position_v4.eps}
				\caption{pattern 10}
				\label{figure:simul1_3_2_j}
			\end{subfigure}
			\caption{Patterns of distribution of users in SIMUL1-3-2.}
			\label{figure:simul1_3_2_p}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul1_3_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul1_3_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul1_3_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul1_3_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul1_3_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul1_3_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul1_3_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul1_3_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul1_3_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp22_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul1_3_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL1-3-2.}
			\label{figure:simul1_3_2_ba}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp22_1-3-2_utility.eps}
				\caption{sum of utility function}
				\label{figure:simul1_3_2_u_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp22_1-3-2_bandwidth.eps}
				\caption{number of users}
				\label{figure:simul1_3_2_u_b}
			\end{subfigure}
			\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL1-3-2.}
			\label{figure:simul1_3_2_u}
		\end{center}
	\end{figure}

\clearpage

%\subsubsection{SIMUL1-4}
	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_01_position_v4.eps}
				\caption{pattern 1}
				\label{figure:simul1_4_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_02_position_v4.eps}
				\caption{pattern 2}
				\label{figure:simul1_4_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_03_position_v4.eps}
				\caption{pattern 3}
				\label{figure:simul1_4_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_04_position_v4.eps}
				\caption{pattern 4}
				\label{figure:simul1_4_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_05_position_v4.eps}
				\caption{pattern 5}
				\label{figure:simul1_4_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_06_position_v4.eps}
				\caption{pattern 6}
				\label{figure:simul1_4_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_07_position_v4.eps}
				\caption{pattern 7}
				\label{figure:simul1_4_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_08_position_v4.eps}
				\caption{pattern 8}
				\label{figure:simul1_4_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_09_position_v4.eps}
				\caption{pattern 9}
				\label{figure:simul1_4_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_10_position_v4.eps}
				\caption{pattern 10}
				\label{figure:simul1_4_j}
			\end{subfigure}
			\caption{Patterns of distribution of users in SIMUL1-4.}
			\label{figure:simul1_4_p}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul1_4_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul1_4_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul1_4_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul1_4_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul1_4_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul1_4_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul1_4_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul1_4_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul1_4_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp23_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul1_4_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL1-4.}
			\label{figure:simul1_4_ba}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp23_1-4_utility.eps}
				\caption{sum of utility function}
				\label{figure:simul1_4_u_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.8\textwidth}
				\centering
				\includegraphics[width=10cm]{fig/exp23_1-4_bandwidth.eps}
				\caption{number of users}
				\label{figure:simul1_4_u_b}
			\end{subfigure}
			\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL1-4.}
			\label{figure:simul1_4_u}
		\end{center}
	\end{figure}

%=====================================
%\subsubsection{SIMUL2-1-1}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul2_1_1_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul2_1_1_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul2_1_1_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul2_1_1_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul2_1_1_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul2_1_1_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul2_1_1_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul2_1_1_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul2_1_1_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul2_1_1_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL2-1-1.}
		\label{figure:simul2_1_1_p}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_01_allocated_bandwidth.eps}
			\caption{bandwidth allocation 1}
			\label{figure:simul2_1_1_b_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_02_allocated_bandwidth.eps}
			\caption{bandwidth allocation 2}
			\label{figure:simul2_1_1_b_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_03_allocated_bandwidth.eps}
			\caption{bandwidth allocation 3}
			\label{figure:simul2_1_1_b_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_04_allocated_bandwidth.eps}
			\caption{bandwidth allocation 4}
			\label{figure:simul2_1_1_b_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_05_allocated_bandwidth.eps}
			\caption{bandwidth allocation 5}
			\label{figure:simul2_1_1_b_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_06_allocated_bandwidth.eps}
			\caption{bandwidth allocation 6}
			\label{figure:simul2_1_1_b_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_07_allocated_bandwidth.eps}
			\caption{bandwidth allocation 7}
			\label{figure:simul2_1_1_b_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_08_allocated_bandwidth.eps}
			\caption{bandwidth allocation 8}
			\label{figure:simul2_1_1_b_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_09_allocated_bandwidth.eps}
			\caption{bandwidth allocation 9}
			\label{figure:simul2_1_1_b_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp10_10_allocated_bandwidth.eps}
			\caption{bandwidth allocation 10}
			\label{figure:simul2_1_1_b_j}
		\end{subfigure}
		\caption{Bandwidth allocation in SIMUL2-1-1.}
		\label{figure:simul2_1_1_ba}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp10_2-1-1_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul2_1_1_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp10_2-1-1_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul2_1_1_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL2-1-1.}
		\label{figure:simul2_1_1_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL2-1-2}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul2_1_2_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul2_1_2_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul2_1_2_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul2_1_2_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul2_1_2_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul2_1_2_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul2_1_2_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul2_1_2_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul2_1_2_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp13_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul2_1_2_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL2-1-2.}
		\label{figure:simul2_1_2_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul2_1_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul2_1_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul2_1_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul2_1_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul2_1_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul2_1_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul2_1_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul2_1_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul2_1_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp13_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul2_1_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL2-1-2.}
			\label{figure:simul2_1_2_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp13_2-1-2_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul2_1_2_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp13_2-1-2_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul2_1_2_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL2-1-2.}
		\label{figure:simul2_1_2_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL2-2-1}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul2_2_1_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul2_2_1_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul2_2_1_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul2_2_1_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul2_2_1_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul2_2_1_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul2_2_1_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul2_2_1_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul2_2_1_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp11_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul2_2_1_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL2-2-1.}
		\label{figure:simul2_2_1_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul2_2_1_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul2_2_1_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul2_2_1_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul2_2_1_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul2_2_1_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul2_2_1_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul2_2_1_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul2_2_1_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul2_2_1_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp11_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul2_2_1_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL2-2-1.}
			\label{figure:simul2_2_1_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp11_2-2-1_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul2_2_1_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp11_2-2-1_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul2_2_1_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL2-2-1.}
		\label{figure:simul2_2_1_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL2-2-2}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul2_2_2_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul2_2_2_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul2_2_2_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul2_2_2_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul2_2_2_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul2_2_2_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul2_2_2_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul2_2_2_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul2_2_2_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp14_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul2_2_2_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL2-2-2.}
		\label{figure:simul2_2_2_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul2_2_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul2_2_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul2_2_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul2_2_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul2_2_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul2_2_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul2_2_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul2_2_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul2_2_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp14_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul2_2_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL2-2-2.}
			\label{figure:simul2_2_2_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp14_2-2-2_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul2_2_2_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp14_2-2-2_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul2_2_2_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL2-2-2.}
		\label{figure:simul2_2_2_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL2-3-1}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul2_3_1_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul2_3_1_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul2_3_1_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul2_3_1_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul2_3_1_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul2_3_1_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul2_3_1_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul2_3_1_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul2_3_1_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp12_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul2_3_1_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL2-3-1.}
		\label{figure:simul2_3_1_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul2_3_1_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul2_3_1_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul2_3_1_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul2_3_1_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul2_3_1_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul2_3_1_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul2_3_1_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul2_3_1_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul2_3_1_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp12_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul2_3_1_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL2-3-1.}
			\label{figure:simul2_3_1_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp12_2-3-1_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul2_3_1_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp12_2-3-1_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul2_3_1_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL2-3-1.}
		\label{figure:simul2_3_1_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL2-3-2}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul2_3_2_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul2_3_2_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul2_3_2_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul2_3_2_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul2_3_2_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul2_3_2_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul2_3_2_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul2_3_2_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul2_3_2_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp15_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul2_3_2_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL2-3-2.}
		\label{figure:simul2_3_2_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul2_3_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul2_3_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul2_3_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul2_3_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul2_3_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul2_3_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul2_3_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul2_3_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul2_3_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp15_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul2_3_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL2-3-2.}
			\label{figure:simul2_3_2_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp15_2-3-2_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul2_3_2_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp15_2-3-2_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul2_3_2_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL2-3-2.}
		\label{figure:simul2_3_2_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL2-4}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul2_4_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul2_4_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul2_4_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul2_4_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul2_4_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul2_4_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul2_4_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul2_4_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul2_4_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp16_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul2_4_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL2-4.}
		\label{figure:simul2_4_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul2_4_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul2_4_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul2_4_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul2_4_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul2_4_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul2_4_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul2_4_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul2_4_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul2_4_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp16_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul2_4_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL2-4.}
			\label{figure:simul2_4_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp16_2-4_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul2_4_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp16_2-4_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul2_4_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL2-4.}
		\label{figure:simul2_4_u}
	\end{center}
\end{figure}

\clearpage

%=====================================
%\subsubsection{SIMUL3-1-1}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul3_1_1_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul3_1_1_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul3_1_1_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul3_1_1_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul3_1_1_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul3_1_1_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul3_1_1_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul3_1_1_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul3_1_1_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp24_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul3_1_1_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL3-1-1.}
		\label{figure:simul3_1_1_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul3_1_1_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul3_1_1_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul3_1_1_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul3_1_1_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul3_1_1_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul3_1_1_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul3_1_1_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul3_1_1_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul3_1_1_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp24_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul3_1_1_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL3-1-1.}
			\label{figure:simul3_1_1_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp24_3-1-1_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul3_1_1_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp24_3-1-1_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul3_1_1_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL3-1-1.}
		\label{figure:simul3_1_1_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL3-1-2}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul3_1_2_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul3_1_2_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul3_1_2_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul3_1_2_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul3_1_2_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul3_1_2_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul3_1_2_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul3_1_2_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul3_1_2_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp27_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul3_1_2_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL3-1-2.}
		\label{figure:simul3_1_2_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul3_1_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul3_1_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul3_1_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul3_1_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul3_1_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul3_1_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul3_1_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul3_1_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul3_1_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp27_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul3_1_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL3-1-2.}
			\label{figure:simul3_1_2_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp27_3-1-2_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul3_1_2_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp27_3-1-2_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul3_1_2_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL3-1-2.}
		\label{figure:simul3_1_2_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL3-2-1}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul3_2_1_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul3_2_1_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul3_2_1_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul3_2_1_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul3_2_1_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul3_2_1_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul3_2_1_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul3_2_1_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul3_2_1_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp25_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul3_2_1_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL3-2-1.}
		\label{figure:simul3_2_1_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul3_2_1_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul3_2_1_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul3_2_1_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul3_2_1_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul3_2_1_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul3_2_1_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul3_2_1_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul3_2_1_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul3_2_1_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp25_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul3_2_1_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL3-2-1.}
			\label{figure:simul3_2_1_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp25_3-2-1_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul3_2_1_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp25_3-2-1_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul3_2_1_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL3-2-1.}
		\label{figure:simul3_2_1_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL3-2-2}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul3_2_2_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul3_2_2_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul3_2_2_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul3_2_2_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul3_2_2_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul3_2_2_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul3_2_2_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul3_2_2_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul3_2_2_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp28_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul3_2_2_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL3-2-2.}
		\label{figure:simul3_2_2_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul3_2_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul3_2_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul3_2_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul3_2_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul3_2_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul3_2_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul3_2_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul3_2_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul3_2_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp28_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul3_2_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL3-2-2.}
			\label{figure:simul3_2_2_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp28_3-2-2_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul3_2_2_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp28_3-2-2_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul3_2_2_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL3-2-2.}
		\label{figure:simul3_2_2_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL3-3-1}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul3_3_1_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul3_3_1_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul3_3_1_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul3_3_1_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul3_3_1_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul3_3_1_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul3_3_1_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul3_3_1_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul3_3_1_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp26_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul3_3_1_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL3-3-1.}
		\label{figure:simul3_3_1_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul3_3_1_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul3_3_1_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul3_3_1_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul3_3_1_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul3_3_1_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul3_3_1_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul3_3_1_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul3_3_1_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul3_3_1_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp26_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul3_3_1_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL3-3-1.}
			\label{figure:simul3_3_1_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp26_3-3-1_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul3_3_1_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp26_3-3-1_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul3_3_1_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL3-3-1.}
		\label{figure:simul3_3_1_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL3-3-2}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul3_3_2_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul3_3_2_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul3_3_2_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul3_3_2_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul3_3_2_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul3_3_2_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul3_3_2_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul3_3_2_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul3_3_2_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp29_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul3_3_2_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL3-3-2.}
		\label{figure:simul3_3_2_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul3_3_2_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul3_3_2_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul3_3_2_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul3_3_2_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul3_3_2_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul3_3_2_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul3_3_2_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul3_3_2_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul3_3_2_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp29_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul3_3_2_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL3-3-2.}
			\label{figure:simul3_3_2_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp29_3-3-2_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul3_3_2_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp29_3-3-2_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul3_3_2_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL3-3-2.}
		\label{figure:simul3_3_2_u}
	\end{center}
\end{figure}

\clearpage

%\subsubsection{SIMUL3-4}
\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_01_position_v4.eps}
			\caption{pattern 1}
			\label{figure:simul3_4_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_02_position_v4.eps}
			\caption{pattern 2}
			\label{figure:simul3_4_b}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_03_position_v4.eps}
			\caption{pattern 3}
			\label{figure:simul3_4_c}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_04_position_v4.eps}
			\caption{pattern 4}
			\label{figure:simul3_4_d}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_05_position_v4.eps}
			\caption{pattern 5}
			\label{figure:simul3_4_e}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_06_position_v4.eps}
			\caption{pattern 6}
			\label{figure:simul3_4_f}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_07_position_v4.eps}
			\caption{pattern 7}
			\label{figure:simul3_4_g}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_08_position_v4.eps}
			\caption{pattern 8}
			\label{figure:simul3_4_h}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_09_position_v4.eps}
			\caption{pattern 9}
			\label{figure:simul3_4_i}
		\end{subfigure}
		\begin{subfigure}[b]{0.32\textwidth}
			\includegraphics[width=\textwidth]{fig/exp30_10_position_v4.eps}
			\caption{pattern 10}
			\label{figure:simul3_4_j}
		\end{subfigure}
		\caption{Patterns of distribution of users in SIMUL3-4.}
		\label{figure:simul3_4_p}
	\end{center}
\end{figure}

	\begin{figure}
		\begin{center}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_01_allocated_bandwidth.eps}
				\caption{bandwidth allocation 1}
				\label{figure:simul3_4_b_a}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_02_allocated_bandwidth.eps}
				\caption{bandwidth allocation 2}
				\label{figure:simul3_4_b_b}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_03_allocated_bandwidth.eps}
				\caption{bandwidth allocation 3}
				\label{figure:simul3_4_b_c}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_04_allocated_bandwidth.eps}
				\caption{bandwidth allocation 4}
				\label{figure:simul3_4_b_d}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_05_allocated_bandwidth.eps}
				\caption{bandwidth allocation 5}
				\label{figure:simul3_4_b_e}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_06_allocated_bandwidth.eps}
				\caption{bandwidth allocation 6}
				\label{figure:simul3_4_b_f}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_07_allocated_bandwidth.eps}
				\caption{bandwidth allocation 7}
				\label{figure:simul3_4_b_g}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_08_allocated_bandwidth.eps}
				\caption{bandwidth allocation 8}
				\label{figure:simul3_4_b_h}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_09_allocated_bandwidth.eps}
				\caption{bandwidth allocation 9}
				\label{figure:simul3_4_b_i}
			\end{subfigure}
			\begin{subfigure}[b]{0.32\textwidth}
				\includegraphics[width=\textwidth]{fig/exp30_10_allocated_bandwidth.eps}
				\caption{bandwidth allocation 10}
				\label{figure:simul3_4_b_j}
			\end{subfigure}
			\caption{Bandwidth allocation in SIMUL3-4.}
			\label{figure:simul3_4_ba}
		\end{center}
	\end{figure}

\begin{figure}
	\begin{center}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp30_3-4_utility.eps}
			\caption{sum of utility function}
			\label{figure:simul3_4_u_a}
		\end{subfigure}
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[width=10cm]{fig/exp30_3-4_bandwidth.eps}
			\caption{number of users}
			\label{figure:simul3_4_u_b}
		\end{subfigure}
		\caption{Comparison of sum of utility function and number of users obtaining more bandwidth in SIMUL3-4.}
		\label{figure:simul3_4_u}
	\end{center}
\end{figure}

\clearpage